---
title: "Statistical Inference Project 2 - Exploring Tooth Growth"
author: "Sukdev Koner"
date: "Sept 24, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
## Overview
In this project, I am going to explore `TootGrowth` data of the R 'datasets' package by comparing the guinea pigs' tooth growth by supplement and dose. First, I would do some exploratory analysis on the data. Then I will use Hypothesis test to make conclusions about what influences the tooth growth.
 
The `ToothGrowth` data set consists of 60 observations and 3 variables:  
1. `len`: Teeth length in millimeters (numeric variable)  
2. `supp`: Supplement type (factor variable with levels VC and OJ)  
3. `dose`: Dose in milligrams (numeric variable) [0.5, 1, and  2 mg] 



### Load ToothGrowth data and perform basic exploratory analysis
```{r echo=FALSE}
library(ggplot2)
data(ToothGrowth)
str(ToothGrowth)
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
toothdata <- ToothGrowth
```

```{r}
summary(ToothGrowth)
```
Here `dose` is a numeric variable but we have to convert to as **Factor** for subsequent inference analysis. 
Here we can see that for 'Supplement Type', it has 2 different levels and 'Dose Level' has three diferent levels.

#### A. Analysis of Tooth Length by Supplement Type :  
```{r fig.height= 2, fig.width= 4}
ggplot(aes(x = supp, y = len), data = toothdata) + 
geom_boxplot(aes(fill = supp)) + xlab("Supplement Type") + 
ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Supplements") + 
scale_fill_manual(name = "Supplement Type", values = c("orange", "brown"), labels = c("Orange juice (OJ)", "Vitamin C (VC)"))
```
This box plot above shows that guinea pigs which received orange juice (OJ) have a greater average tooth length than those pigs which got Vitamin C (VC). 


#### B. Analysis of Tooth Length by Dose  Level :
```{r fig.height= 2, fig.width= 4}
ggplot(aes(x = dose, y = len), data = toothdata) + 
geom_boxplot(aes(fill = dose)) + xlab("Dose [mg]") + 
ylab("Tooth Length [mm]") + ggtitle("Tooth Length by Dose Levels") + 
scale_fill_manual(name = "Dose level", values = c("darkblue", "green", "pink"), labels = c("0.5mg", "1mg", "2mg"))
```
This plot states that when we increase the Dose, we are seeing increase in Average(mean) tooth length. And the 'Range (Max - Min)' for Tooth Length for various 'Doses' are so spread out that they almost did not overlap.

## Hypothesis tests to compare tooth growth by Supplement and Dose 

### 1. By Supplement Type:
We will check if the Difference in 'Tooth length means' between pigs who received Orange Juice and who received Vitamin C, is statistically different from 0.

####Hypothesis
Supplement type has no impact on tooth growth.   
H~0~: Both groups (OJ and VC) have the same mean.  
H~A~: Means are different.  

We will use t-test with unequal variances of two samples. 
```{r}
t <- t.test(len ~ supp, data = toothdata, var.equal = FALSE)
t$conf.int
t$p.value
```
The p-value is greater than Significance Level($\alpha$ = 0.05) and "Confidence Interval" of the test contains zero,so we **fail to reject Null Hypothesis**. Hence we can say that **'Supplement Types' seems to have no impact on Tooth growth.**


### 2. By Dose Level:
We will check if the Difference in 'Tooth length means' between pigs who received different Dose is statistically different from 0. We will first compare between **dose 2mg and dose 0.5mg**.
After that, we will see the comparison test between **dose 1mg and dose 0.5mg** and then between **dose 2mg and dose 1mg**.

####Hypothesis
Higher doses of supplement causes less tooth growth.  
H~0~: Mean of Dose level 2 is smaller or equal than dose level 0.5.    
H~A~: Mean of level 2 is greater than level 0.5.   
```{r}
# run t-test using dose amounts 0.5 and 2.0
t1 <- t.test(len~dose,data= subset(toothdata, toothdata$dose %in% c(0.5,2.0)))
Dose_half_Vs_2_Mg <- c(round(t1$conf.int,3), round(t1$p.value,15))
# run t-test using dose amounts 1.0 and 0.5
t2 <- t.test(len~dose,data= subset(toothdata, toothdata$dose %in% c(1.0,0.5)))
Test_type <- c("Conf. Interval-Low", "Conf. Interval-High","P-Value")
Dose_1_Vs_half_Mg <- c(round(t2$conf.int,3), round(t2$p.value,9))
# run t-test using dose amounts 2.0 and 1.0
t3 <- t.test(len~dose,data= subset(toothdata, toothdata$dose %in% c(2.0,1.0)))
Dose_2_Vs_1_Mg <- c(round(t3$conf.int,3), round(t3$p.value,9))
r1 <- rbind(Test_type,Dose_half_Vs_2_Mg,Dose_1_Vs_half_Mg,Dose_2_Vs_1_Mg)
r1
```
After comparing all results, their p-values are very low (less than significance level) and the confidence intervals also do not contain zero, so we can **reject the null hypothesis** and conclude that **'Dose' does affect the tooth length**.

### Conclusion
According the t-tests, we can conclude that  
- **Supplement type** has no impact of thooth growth  
- increasing **Dose Level** results in increasing tooth growth    

### Assumptions
Following assumptions are made here:  
 - each subject is randomly assigned and representative of population  
 - variables are independent and identically distributed (i.i.d.).  
 
[Note : The complete assignment can be found on [GitHub](https://github.com/sukdevkoner/datasciencecoursera/tree/master/Course6_Statistical_Inference).]
